Python as a machine learning language: experiences from the scikit-learn project

### 2013-05-09 paper.tex
Small typo and style changes

In the remainder of the paper, we use the following notation. Training data are
denoted by \Xtr ~and test data (i.e., unseen data to which the learning
algorithm need be able to generalize) by \Xte.  Following the discussion above,
we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
such as classification or regression, we denote the training and test labels by
\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
sample's label value).

@@

The bulk of the scikit-learn application programming interface
consists of so-called \textit{estimator} objects.
All learning algorithms for classification, regression and clustering
are offered as such objects,
but so are feature extraction, feature selection and dimensionality reduction.

@@

Importantly, the parameters learned by an estimator
are exposed as \textit{public} attributes on the trained object.
This facilitates model inspection
and makes it possible to train a model using scikit-learn,
export it to a (standardized or custom) external data format
and use the learned parameters in a different piece of code,
perhaps a prediction algorithm implemented in a different language
or one that combines the model parameters with other information
to build a custom model.
Of course, it also forces the developers to think carefully
about what might in a more ``black box''-style toolkit
be considered the internals of objects,
since authors are client code are allowed to rely
on the names and formats of model parameters.

A \texttt{fit} method is present even on objects
that perform seemingly mundane tasks such as vectorizing text documents
for subsequent learning.
Such classes, however, can be said to learn their vocabulary
from the training corpus,
and follow the conventions set out above.
Even those estimators that are entirely stateless and therefore
do not require a \texttt{fit} method to perform useful work,
nevertheless have such a method for consistency and composibility,
and to perform parameter validation.
Examples of this latter kind of objects are the normalizing transformer,
kernel approximation transformers
\citep{rahimi2007random, li2010random, vedaldi2010efficient},
and two feature extraction classes that implement the ``hashing trick''
\citep{weinberger2009}.

@@

A distinguishing feature of scikit-learn estimators
is that they can be composed into \texttt{Pipeline} objects.
Not only do such objects combine typical machine learning workflows
(extracting features, centering/scaling/normalization,
feature selection, making predictions)
in a single object that is itself an estimator,
and whose methods cause data to flow through the pipeline.
More importantly, though, they can be used for
systematic, automated model selection.
This is best explained by example.

By composing, say, a \texttt{CountVectorizer} object
(which extracts term frequency features from text documents
according to a bag-words model),
a \textsf{tf--idf} transformer, a $\chi^2$ feature selector and a linear SVM,
one obtains a pipeline object with various hyperparameters:
whether to do stopword filtering,
linear vs.\ logarithmic \textsf{tf}, number of features to keep,
SVM regularization ($L_1$ or $L_2$ norm, strength $C$) a.o.
Such a composite \texttt{Pipeline} object
exposes all of these parameters using a special syntax;
each estimators must be given a name,
and if the SVM's name is, e.g., \texttt{linearsvm},
then its $C$ parameter is exposed as \texttt{linearsvm\_\_C}.
The pipeline's \texttt{set\_params} splits the parameter names
on double underscores, uses the first part of the result
to look up the estimator,
and passes the second part to the estimator's own \texttt{set\_params}
as a keyword argument; due to Python's dynamic parameter passing,
the result is the same as if the user had called \texttt{set\_params}
on the SVM themselves.
(The double underscore syntax is not special to pipelines;
it is also employed in ensemble estimators such as random forests.)

The aforementioned \texttt{Pipeline} object
can be passed as an argument to a model selection algorithm.
Available algorithms at the time of writing include random search
\citep{bergstra2012} and ``grid search''.
The latter, which we shall use in this example,
performs an exhaustive search through a grid formed by the Cartesian product
of a set of possible values for each parameter, specified by the user.
In our example, this grid might be given as

@@

At each point in this grid, $k$-fold cross validation is run
to estimate the performance of the estimator according to some measure
(e.g.\ accuracy, $F_\beta$-score or a host of clustering quality metrics)
and the best performing set of hyperparameters is stored
on the \texttt{GridSearchCV} object; again, as a public attribute.

@@

scikit-learn is primarily implemented in Python and Cython
\citep{behnel2011cython},
a language that extends Python with static typing
and a compiler that produces C extension modules
for the Python runtime system.
In addition, it includes (modified versions of)
the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
used for training support vector machines
and logistic regression models \citep{chang2011libsvm, fan2008}.
These libraries, both written in C{}\verb!++!,
and wrapped using Cython modules.

>>>>

In the remainder of the paper, we use the following notation. Training data are
denoted by \Xtr ~and test data (i.e., unseen data to which the learning
algorithm needs to be able to generalize) by \Xte.  Following the discussion above,
we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
such as classification or regression, we denote the training and test labels by
\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
sample's label value).

@@

The bulk of the scikit-learn application programming interface
consists of so-called \textit{estimator} objects.
All learning algorithms for classification, regression and clustering
are offered as such objects,
but so are feature extraction, feature selection and dimensionality reduction
algorithms.

@@

Importantly, the parameters learned by an estimator
are exposed as \textit{public} attributes on the trained object.
This facilitates model inspection
and makes it possible to train a model using scikit-learn,
export it to a (standardized or custom) external data format
and use the learned parameters in a different piece of code,
perhaps a prediction algorithm implemented in a different language
or one that combines the model parameters with other information
to build a custom model.
% XXX Hard to follow phrasing below
Of course, it also forces the developers to think carefully
about what might, in a more ``black box''-style toolkit,
be considered the internals of objects,
since authors of client code are allowed to rely
on the names and formats of model parameters.

A \texttt{fit} method is present even on objects
that perform seemingly mundane tasks such as vectorizing text documents
for subsequent learning.
Such classes, however, can be said to learn their vocabulary
from the training corpus,
and follow the conventions set out above.
Even those estimators that are entirely stateless and therefore
do not require a \texttt{fit} method to perform useful work,
nevertheless have such a method for consistency and composibility,
and to perform parameter validation.
Examples of this latter kind of objects are the normalizing transformer,
kernel approximation transformers
\citep{rahimi2007random, li2010random, vedaldi2010efficient},
% XXX is it relevant that they are exactly two? Propose s/two/the
and two feature extraction classes that implement the ``hashing trick''
\citep{weinberger2009}.

@@

A distinguishing feature of scikit-learn estimators
is that they can be composed into \texttt{Pipeline} objects.
Such objects combine typical machine learning workflows
(extracting features, centering/scaling/normalization,
feature selection, making predictions)
in a single object that is itself an estimator,
and whose methods cause data to flow through the pipeline.
More importantly, though, they can be used for
systematic, automated model selection.
This is best explained by example.

By composing, say, a \texttt{CountVectorizer} object
(which extracts term frequency features from text documents
according to a bag-words model),
a \textsf{tf--idf} transformer, a $\chi^2$ feature selector and a linear SVM,
one obtains a pipeline object with various hyperparameters:
whether to do stopword filtering,
linear vs.\ logarithmic \textsf{tf}, number of features to keep,
SVM regularization ($L_1$ or $L_2$ norm, strength $C$) a.o.
Such a composite \texttt{Pipeline} object
exposes all of these parameters using a special syntax;
each estimators must be given a name,
and if the SVM's name is, e.g., \texttt{linearsvm},
then its $C$ parameter is exposed as \texttt{linearsvm\_\_C}.
The pipeline's \texttt{set\_params} splits the parameter names
on double underscores, uses the first part of the result
to look up the estimator,
and passes the second part to the estimator's own \texttt{set\_params}
as a keyword argument; due to Python's dynamic parameter passing,
the result is the same as if the user had called \texttt{set\_params}
on the SVM themselves.
(The double underscore syntax is not special to pipelines;
it is also employed in ensemble estimators such as random forests
% Is the following clear? Suggestions to replace "direct attributes" with"?
It is chosen as a convention to avoid ambiguity, as direct
attributes and methods of scikit-learn objects never contain
double underscores.)

The aforementioned \texttt{Pipeline} object
can be passed as an argument to a model selection algorithm.
Available algorithms at the time of writing include random search
\citep{bergstra2012} and ``grid search''.
The latter, which we shall use in this example,
performs an exhaustive search through a grid formed by the Cartesian product
of a set of possible values for each parameter, specified by the user.
In our example, this grid might be given as:

@@

At each point in this grid, $k$-fold cross validation is run
to estimate the performance of the estimator according to some measure
(e.g.\ accuracy, $F_\beta$-score or a host of clustering quality metrics)
and the best performing set of hyperparameters is stored
on the \texttt{GridSearchCV} object, again, as a public attribute.

@@

scikit-learn is primarily implemented in Python and Cython
\citep{behnel2011cython},
a language that extends Python with static typing
and a compiler that produces C extension modules
for the Python runtime system.
In addition, it includes (modified versions of)
the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
used for training support vector machines
and logistic regression models \citep{chang2011libsvm, fan2008}.
These libraries are both written in C{}\verb!++!
and wrapped using Cython modules.
